```{r include = FALSE}
if(!knitr:::is_html_output())
{
  options("width"=56)
  knitr::opts_chunk$set(tidy.opts=list(width.cutoff=56, indent = 2), tidy = TRUE)
  knitr::opts_chunk$set(fig.pos = 'H')
}
```

# What did we accomplish?

I have said this earlier in this book, my initial goal for meeting with a trainer and going to the gym was to lose weight, but as we started sweating it out I realized that we were getting a lot more out of this exercise and clean eating regimen than just a lighter body. Even so, weight loss and overall getting the body in a better shape are important goals, so how did we fare on these?


```{r echo=FALSE, warning=FALSE, message=FALSE}
# we want to create a dataframe which contain multiple different metrics which can all
  # be displayed in one single table. We do this by creating a single row dataframe for each
  # metric of interest and then row binding (concatenating) each of these individual dataframes.
    
  # starting weight. Since the weight loss dataframe is sorted
  # in asc order of the date so the first date for each group (name in this case)
  # is automatically the row that contains the starting weight. This way
  # we have a two row dataframe which we can spread to convert to a single row dataframe
  # such that the group i.e. name of person becomes a column name and then we add a new column
  # saying this is the metric that this single row dataframe is tracking. This same strategy
  # of using spread and adding a metric column is followed for all individual dataframes.
  df_starting_wt <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==1) %>%
    mutate(value=as.character(value)) %>%
    select(name, value) %>%
    spread(name, value) %>%
    mutate(Metric="Starting Weight (lb)") %>%
    select(Metric, everything())
  
  # current weight dataframe  
  df_curr_wt <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==max(rn)) %>%
    mutate(value=as.character(value)) %>%
    select(name, value) %>%
    spread(name, value) %>%
    mutate(Metric="Current Weight (lb)") %>%
    select(Metric, everything())
  
  # weight loss summary i.e. how much have we lost  
  df_wt_loss_summary <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==1 | rn==max(rn)) %>%
    mutate(total_wt_loss = as.character(-1*(value-lag(value)))) %>%
    drop_na() %>%
    select(name, total_wt_loss) %>%
    spread(name, total_wt_loss) %>%
    mutate(Metric="Total weight loss (lb)") %>%
    select(Metric, everything())
  
  # total duration  
  df_duration_summary <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==1 | rn==max(rn)) %>%
    mutate(days_since_start = as.character((Date-lag(Date)))) %>%
    drop_na() %>%
    select(name, days_since_start) %>%
    spread(name, days_since_start) %>%
    mutate(Metric="Days since start") %>%
    select(Metric, everything())
  
  # duration for losing the last 10 pounds
  # this is a tricky one, trying to do everythingin a single dplyr pipeline
  # what we do is as follows: calc the cumulative loss from the latest date
  # and then assign a row number, the rn==1 corresponds to latest date so we
  # want that entry, then we want to find the first entry for which cumul loss >= (say) 10 pounds
  # so we use filter to keep either rn==1 or cumul_loss >= 10, but we only want to keep
  # the first entry where cumul loss beame >=10 so we assign another
  # row number and now we want to the first two rows and then take a diff of the two
  # rows to get number of days. Best to run this outside of shiny to understand.
  df_days_for_last_n_pounds <- df_tidy %>%
    filter(metric=="Weight") %>%
    select(name, Date, value) %>%
    group_by(name) %>%
    arrange(desc(Date), .by_group=TRUE) %>%
    mutate(loss_per_day = value-lag(value, 1)) %>%
    drop_na() %>%
    mutate(cumul_loss = cumsum(loss_per_day)) %>%
    arrange(desc(Date)) %>%
    mutate(rn=row_number()) %>%
    filter(cumul_loss >= N_FOR_LAST_N_POUNDS_OF_INTREST | rn==1) %>%
    mutate(rn2=row_number()) %>%
    filter(rn==1 | rn2==2) %>%
    select(name, Date) %>%
    mutate(days_taken_to_lost_last_n_pounds = as.character(-1*(Date-lag(Date)))) %>%
    drop_na() %>%
    select(name, days_taken_to_lost_last_n_pounds) %>%
    ungroup() %>%
    spread(name, days_taken_to_lost_last_n_pounds) %>%
    mutate(Metric=glue("Days taken to lose last {N_FOR_LAST_N_POUNDS_OF_INTREST} pounds")) %>%
    select(Metric, everything())
  
  # which month did we have the max weight loss (remember weight loss has already been converted to a positive value
  # so loss is +ve and gain is -ve at this point)
  df_best_wt_loss_month <- df_wt_loss %>%
    mutate(m = MONTH_ABB[month(Date)]) %>%
    group_by(name, m) %>%
    summarize(total_wt_loss = sum(loss_per_day)) %>%
    filter(total_wt_loss == max(total_wt_loss, na.rm=TRUE)) %>%
    mutate(best_wt_loss_month = paste0(m, ", ", round(total_wt_loss, 2), " lb")) %>%
    select(name, best_wt_loss_month) %>%
    ungroup() %>%
    spread(name,best_wt_loss_month) %>%
    mutate(Metric="Best Weight loss month") %>%
    select(Metric, everything())
  
  
  # join all the dataframes and display them in the table
  df_imp_metrics <- bind_rows(df_duration_summary,
            df_days_for_last_n_pounds,
            df_starting_wt,
            df_curr_wt,
            df_wt_loss_summary,
            df_best_wt_loss_month)

  gt(df_imp_metrics) %>%
    tab_header(
      title = md("**Important Metrics**"),
      subtitle = md("Key data points that describe the journey")
    )  %>%
    tab_source_note(
      source_note = md(CAPTION)
    ) %>%
    tab_footnote(
      footnote = md(glue("{round(p2_wt_lost_as_pct, 2)}% of the starting body weight.")),
      locations = cells_body(columns = (P2_NAME), rows = 5
    )) %>%
    tab_footnote(
      footnote = md(glue("{round(p1_wt_lost_as_pct, 2)}% of the starting body weight.")),
      locations = cells_body(columns = (P1_NAME), rows = 5
    ))
```

## Percentages are revealing

So net-net in about 7 months, I lost about 20% of my body weight and Nidhi lost about 15%. Not too bad. In terms of how far we have progressed compared to the goals we started with, I have some more miles to go while Nidhi is almost there.

```{r echo=FALSE}
df2 <- data.frame(name = c(P1_NAME, P2_NAME),
                  title = c(P1_NAME, P2_NAME),
                  group = c(ifelse(p1_target_achieved_pct >= 75, "green", "orange"),
                            ifelse(p2_target_achieved_pct >= 75, "green", "orange")),
                  percentage = c(p1_target_achieved_pct/100, p2_target_achieved_pct/100),
                  label = c(glue("{round(p1_target_achieved_pct, 2)}%"),
                            glue("{round(p2_target_achieved_pct, 2)}%")))
                  
ggplot(df2, aes(fill = group, ymax = percentage, ymin = 0, xmax = 2, xmin = 1)) +
  geom_rect(aes(ymax=1, ymin=0, xmax=2, xmin=1), fill ="#ece8bd") +
  geom_rect() + 
  coord_polar(theta = "y",start=-pi/2) + xlim(c(0, 2)) + ylim(c(0,2)) +
  #geom_text(aes(x = 0, y = 0, label = label, colour=group), size=6.5, family="Poppins SemiBold") +
  #geom_text(aes(x=1.5, y=1.5, label=title), family="Poppins Light", size=4.2) + 
  geom_text(aes(x = 0, y = 0, label = label, colour=group), size=6.5) +
  geom_text(aes(x=1.5, y=1.5, label=title), size=4.2) + 
  facet_wrap(~title, ncol = 5) +
  theme_void() +
  scale_fill_tableau() + 
  #scale_fill_manual(values = c("red"="#C9146C", "orange"="#DA9112", "green"="#129188")) +
  #scale_colour_manual(values = c("red"="#C9146C", "orange"="#DA9112", "green"="#129188")) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank()) +
  guides(fill=FALSE) +
  guides(colour=FALSE) + 
  labs(title=glue("How much of the weight loss goal has been accomplished as of {max(df_tidy$Date, na.rm=TRUE)}?"))

```

## Changes in other biometrics

Along with the body weight, other metrics also saw significant change. This is seen in the following charts. BMI is widely used (I suppose accepted as well) measurement to determine if someone is healthy, obese or overweight. We saw reduction in BMI as well which as expected is correlated to the reduction in weight. NIH guidelines for BMI are available [here](https://www.nhlbi.nih.gov/health/educational/healthdisp/pdf/tipsheets/Are-You-at-a-Healthy-Weight.pdf) for reference.

All measurements in the charts below were done automatically as part of the daily metrics measured by the scale and synched with our phone. This made it really easy to collect and analyze this data.

```{r echo=FALSE, fig.height=8, fig.width=12}
  df_tidy %>%
  ggplot(aes(x=Date, y=value, col=name)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  facet_wrap(~metric, scales = "free_y") + 
  theme_fivethirtyeight() +
  xlab("") +
  labs(title=glue(MAIN_PAGE_CHART_TITLE),
       subtitle=glue("Tracking weight and other biometrics, Timespan: {min(df_tidy$Date)} to {max(df_tidy$Date)}"),
       caption=CAPTION) + 
  theme(text = element_text(size=CHART_ELEMENT_TEXT_SIZE-5), legend.title = element_blank()) + 
  scale_color_tableau()
```

## Body measurements

Weight and BMI is not the only metric we tracked, these metrics are often times all we think in terms of measuring but there is more. Just as the proof of the pudding is in the eating, the proof of the weight loss is in the wearing (of clothes). As we kept on making progress in our journey, the clothes started fitting well at first, then getting loose and the finally it reached a point where most of our old clothes were just too big and this necessitated a wardrobe refresh. A happy problem to have.

We tracked this by doing body measurements once every few months. The charts below have a story to tell.

```{r echo=FALSE, fig.height=8, fig.width=12}
df_measurements %>%
    ggplot(aes(x=date, y=value, col=name)) +
    geom_line() + 
    geom_point(show.legend = FALSE) + 
    geom_text(aes(label = paste0("(", value, ")")), nudge_y = -0.015*df_measurements$value, show.legend = FALSE) + 
    labs(title="Changes in body measurements with diet and excercise",
         subtitle=glue("All measurements in inches. Timespan: {min(df_measurements$date)} to {max(df_measurements$date)}"),
         caption="Data Source: measurements done @home") +
    facet_wrap(~measurement, scales="free") +
    theme_fivethirtyeight() + 
    scale_color_tableau() + 
    theme(text = element_text(size=CHART_ELEMENT_TEXT_SIZE-5), legend.title = element_blank())
```

## A forecast and a promise

As we were going through this journey, I was very eager to apply some forecasting and determine if we could project a reasonable date when we would be able to meet our weight loss target. As much as this book is not just about weight loss, there is no denying the fact that it was one of the most (if not the most) tangible outcome we were tracking towards.

Once we had collected a reasonable amount of data, I used standard timeseries forecasting techniques to determine how our weights would look say 30, 90 or 180 days from the current date. I used the [Prophet](https://facebook.github.io/prophet/#:~:text=Prophet%20is%20a%20procedure%20for,daily%20seasonality%2C%20plus%20holiday%20effects.&text=Prophet%20is%20robust%20to%20missing,and%20typically%20handles%20outliers%20well.) package from Facebook AI Research ([FAIR](https://github.com/facebookresearch)) to do the timeseries forecast. The results are presented below. This forecast was done on August 16, 2020 and as per this forecast we should be able to achieve our target weight in early October for Nidhi and mid-December for me.

Once we had these forecasted date and the plots created, we started monitoring very closely if our daily weight measurements were within the range of errors as shown in these plots. Some days the weight did creep out of the error limits but then it served as a nice tool to keep us honest, so in a manner of speaking we knew how much we could deviate. So if an Indian takeout dinner one evening did set us back, we knew we had to make up for it in the next few days to come back within acceptable range. So far it seems the projects are holding up reasonably well. If nothing else, it provides certain guard rails to not let one go completely off track. The promise still is, if we keep doing what we have been diligently doing, we should be able to achieve our goals around, if not exactly on, the forecasted date. It is important to mention here, this simple forecast is considering the weight loss as a univariate timeseries i.e. single variable timeseries, we know it is much more complicated than that. The premise is that all exogenous factors remaining the same, this forecast should hold up.

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=12}
df_forecast_p2 %>%
  rename(Actual=y, Forecast='yhat') %>%
  gather(category, value, -ds, -yhat_lower, -yhat_upper) %>%
  ggplot(aes(x=ds, y=value, col=category)) +
  geom_ribbon(aes(ymin=yhat_lower, ymax=yhat_upper), alpha=.0625, show.legend = FALSE, col=I("lightblue"), fill=I("lightblue")) +
  geom_line() + 
  annotate("point", x = df_target_achieved_p2$date, y = df_target_achieved_p2$target, colour = "darkgreen", size = 1.5, alpha=1) +
  annotate(
    geom = "curve", xend = df_target_achieved_p2$date, yend = df_target_achieved_p2$target, x = df_target_achieved_p2$date+2, y = df_target_achieved_p2$target+2, 
    curvature = .3, arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(geom = "text", x = df_target_achieved_p2$date+2, y = df_target_achieved_p2$target+2, label = glue("Target: {df_target_achieved_p2$target} lb, Date: {df_target_achieved_p2$date}"), hjust = "left") +
  scale_color_tableau() + 
  scale_fill_tableau() + 
  theme_fivethirtyeight() +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y") +
  scale_y_continuous(breaks = seq(180, 260, by = 10)) + 
  xlab("") +
  labs(title=glue("When will Amit achieve his target weight?"),
       subtitle=glue("Light blue lines show lower and upper ranges for forecast. Forecast done on {FORECAST_DATE}")) +
  theme(text = element_text(size=CHART_ELEMENT_TEXT_SIZE-5), legend.title = element_blank()) +
  theme(axis.title = element_text()) + ylab('Weight (lb)')  


```
  
  
  
```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=8, fig.width=12}
df_forecast_p1 %>%
  rename(Actual=y, Forecast='yhat') %>%
  gather(category, value, -ds, -yhat_lower, -yhat_upper) %>%
  ggplot(aes(x=ds, y=value, col=category)) +
  geom_ribbon(aes(ymin=yhat_lower, ymax=yhat_upper), alpha=.0625, show.legend = FALSE, col=I("lightblue"), fill=I("lightblue")) +
  geom_line() + 
  annotate("point", x = df_target_achieved_p1$date, y = df_target_achieved_p1$target, colour = "darkgreen", size = 1.5, alpha=1) +
  annotate(
    geom = "curve", xend = df_target_achieved_p1$date, yend = df_target_achieved_p1$target, x = df_target_achieved_p1$date+2, y = df_target_achieved_p1$target+2, 
    curvature = .3, arrow = arrow(length = unit(2, "mm"))
  ) +
  annotate(geom = "text", x = df_target_achieved_p1$date+2, y = df_target_achieved_p1$target+2, label = glue("Target: {df_target_achieved_p1$target} lb, Date: {df_target_achieved_p1$date}"), hjust = "left") +
  scale_color_tableau() + 
  scale_fill_tableau() + 
  theme_fivethirtyeight() +
  scale_x_date(date_breaks = "3 month", date_labels = "%b %Y") +
  scale_y_continuous(breaks = seq(100, 160, by = 5)) + 
  xlab("") +
  labs(title=glue("When will Nidhi achieve her target weight?"),
       subtitle=glue("Light blue lines show lower and upper ranges for forecast. Forecast done on {FORECAST_DATE}")) +
  theme(text = element_text(size=CHART_ELEMENT_TEXT_SIZE-5), legend.title = element_blank()) +
  theme(axis.title = element_text()) + ylab('Weight (lb)')  

```

