# What did we accomplish?

I have said this earlier in this book, my initial goal for meeting with a trainer and going to the gym was to lose weight, but as we started sweating it out I realized that we were getting a lot more out of this exercise and clean eating regimen than just a lighter body. Even so, weight loss and overall getting the bad in a better shape are important goals, so how did we fare on these?


```{r echo=FALSE}
# we want to create a dataframe which contain multiple different metrics which can all
  # be displayed in one single table. We do this by creating a single row dataframe for each
  # metric of interest and then row binding (concatenating) each of these individual dataframes.
    
  # starting weight. Since the weight loss dataframe is sorted
  # in asc order of the date so the first date for each group (name in this case)
  # is automatically the row that contains the starting weight. This way
  # we have a two row dataframe which we can spread to convert to a single row dataframe
  # such that the group i.e. name of person becomes a column name and then we add a new column
  # saying this is the metric that this single row dataframe is tracking. This same strategy
  # of using spread and adding a metric column is followed for all individual dataframes.
  df_starting_wt <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==1) %>%
    mutate(value=as.character(value)) %>%
    select(name, value) %>%
    spread(name, value) %>%
    mutate(Metric="Starting Weight (lb)") %>%
    select(Metric, everything())
  
  # current weight dataframe  
  df_curr_wt <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==max(rn)) %>%
    mutate(value=as.character(value)) %>%
    select(name, value) %>%
    spread(name, value) %>%
    mutate(Metric="Current Weight (lb)") %>%
    select(Metric, everything())
  
  # weight loss summary i.e. how much have we lost  
  df_wt_loss_summary <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==1 | rn==max(rn)) %>%
    mutate(total_wt_loss = as.character(-1*(value-lag(value)))) %>%
    drop_na() %>%
    select(name, total_wt_loss) %>%
    spread(name, total_wt_loss) %>%
    mutate(Metric="Total weight loss (lb)") %>%
    select(Metric, everything())
  
  # total duration  
  df_duration_summary <- df_wt_loss %>%
    mutate(rn=row_number()) %>%
    filter(rn==1 | rn==max(rn)) %>%
    mutate(days_since_start = as.character((Date-lag(Date)))) %>%
    drop_na() %>%
    select(name, days_since_start) %>%
    spread(name, days_since_start) %>%
    mutate(Metric="Days since start") %>%
    select(Metric, everything())
  
  # duration for losing the last 10 pounds
  # this is a tricky one, trying to do everythingin a single dplyr pipeline
  # what we do is as follows: calc the cumulative loss from the latest date
  # and then assign a row number, the rn==1 corresponds to latest date so we
  # want that entry, then we want to find the first entry for which cumul loss >= (say) 10 pounds
  # so we use filter to keep either rn==1 or cumul_loss >= 10, but we only want to keep
  # the first entry where cumul loss beame >=10 so we assign another
  # row number and now we want to the first two rows and then take a diff of the two
  # rows to get number of days. Best to run this outside of shiny to understand.
  df_days_for_last_n_pounds <- df_tidy %>%
    filter(metric=="Weight") %>%
    select(name, Date, value) %>%
    group_by(name) %>%
    arrange(desc(Date), .by_group=TRUE) %>%
    mutate(loss_per_day = value-lag(value, 1)) %>%
    drop_na() %>%
    mutate(cumul_loss = cumsum(loss_per_day)) %>%
    arrange(desc(Date)) %>%
    mutate(rn=row_number()) %>%
    filter(cumul_loss >= N_FOR_LAST_N_POUNDS_OF_INTREST | rn==1) %>%
    mutate(rn2=row_number()) %>%
    filter(rn==1 | rn2==2) %>%
    select(name, Date) %>%
    mutate(days_taken_to_lost_last_n_pounds = as.character(-1*(Date-lag(Date)))) %>%
    drop_na() %>%
    select(name, days_taken_to_lost_last_n_pounds) %>%
    ungroup() %>%
    spread(name, days_taken_to_lost_last_n_pounds) %>%
    mutate(Metric=glue("Days taken to lose last {N_FOR_LAST_N_POUNDS_OF_INTREST} pounds")) %>%
    select(Metric, everything())
  
  # which month did we have the max weight loss (remember weight loss has already been converted to a positive value
  # so loss is +ve and gain is -ve at this point)
  df_best_wt_loss_month <- df_wt_loss %>%
    mutate(m = MONTH_ABB[month(Date)]) %>%
    group_by(name, m) %>%
    summarize(total_wt_loss = sum(loss_per_day)) %>%
    filter(total_wt_loss == max(total_wt_loss, na.rm=TRUE)) %>%
    mutate(best_wt_loss_month = paste0(m, ", ", round(total_wt_loss, 2), " lb")) %>%
    select(name, best_wt_loss_month) %>%
    ungroup() %>%
    spread(name,best_wt_loss_month) %>%
    mutate(Metric="Best Weight loss month") %>%
    select(Metric, everything())
  
  
  # join all the dataframes and display them in the table
  df_imp_metrics <- bind_rows(df_duration_summary,
            df_days_for_last_n_pounds,
            df_starting_wt,
            df_curr_wt,
            df_wt_loss_summary,
            df_best_wt_loss_month)

  gt(df_imp_metrics) %>%
    tab_header(
      title = md("**Important Metrics**"),
      subtitle = md("Key data points that describe the journey")
    )  %>%
    tab_source_note(
      source_note = md(CAPTION)
    ) %>%
    tab_footnote(
      footnote = md(glue("{round(p2_wt_lost_as_pct, 2)}% of the starting body weight.")),
      locations = cells_body(columns = (P2_NAME), rows = 5
    )) %>%
    tab_footnote(
      footnote = md(glue("{round(p1_wt_lost_as_pct, 2)}% of the starting body weight.")),
      locations = cells_body(columns = (P1_NAME), rows = 5
    ))
```

So net-net in about 7 months, I lost about 20% of my body weight and Nidhi lost about 15%. Not too bad. In terms of how far we have progressed compared to the goals we started with, I have some more miles to go while Nidhi is almost there.

```{r}
df2 <- data.frame(name = c(P1_NAME, P2_NAME),
                  title = c(P1_NAME, P2_NAME),
                  group = c(ifelse(p1_target_achieved_pct >= 75, "green", "orange"),
                            ifelse(p2_target_achieved_pct >= 75, "green", "orange")),
                  percentage = c(p1_target_achieved_pct/100, p2_target_achieved_pct/100),
                  label = c(glue("{round(p1_target_achieved_pct, 2)}%"),
                            glue("{round(p2_target_achieved_pct, 2)}%")))
                  
ggplot(df2, aes(fill = group, ymax = percentage, ymin = 0, xmax = 2, xmin = 1)) +
  geom_rect(aes(ymax=1, ymin=0, xmax=2, xmin=1), fill ="#ece8bd") +
  geom_rect() + 
  coord_polar(theta = "y",start=-pi/2) + xlim(c(0, 2)) + ylim(c(0,2)) +
  geom_text(aes(x = 0, y = 0, label = label, colour=group), size=6.5, family="Poppins SemiBold") +
  geom_text(aes(x=1.5, y=1.5, label=title), family="Poppins Light", size=4.2) + 
  facet_wrap(~title, ncol = 5) +
  theme_void() +
  scale_fill_tableau() + 
  #scale_fill_manual(values = c("red"="#C9146C", "orange"="#DA9112", "green"="#129188")) +
  #scale_colour_manual(values = c("red"="#C9146C", "orange"="#DA9112", "green"="#129188")) +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank()) +
  guides(fill=FALSE) +
  guides(colour=FALSE) + 
  labs(title=glue("How much of the weight loss goal has been accomplished as of {max(df_tidy$Date, na.rm=TRUE)}?"))

```