# From here, where?

```{r echo=FALSE}
library(jpeg)
library(grid)
img <- readJPEG("pictures/before_and_after.jpg")
grid.raster(img)
```

```{r echo=FALSE}
library(jpeg)
library(grid)
img <- readJPEG("pictures/sep6_2021.jpg")
grid.raster(img)
```

Here are the customary before, after and the not  so customary the year after pictures. I am not a particularly picture friendly person but I do think a picture here highlights the amount of change much more so than numbers ever can.

The good thing about journaling (or writing a book for that matter) is that when you go back and revisit the written word it immediately transports you back in time and then hindsight makes you realize "oh, that is what I was thinking, how naive was I?". What I think now about the road ahead is different from what I was thinking same time last year which is what was documented in the first edition of this book. Last year I was thinking purely in terms of fitness goals but now my field of view is much wider. Anyhow, I am keeping the section from the first edition as is and adding a new one for the second edition as well.

## The road ahead, the 2021 version

While 2 years of work cannot bring back 2 lost decades of very little work on my physical and mental health, but the human body is still a very resilient machine. As I write at another place in this book "the right time to start was yesterday, the best time to start is today!". 

I want to extend my new found levels of physical fitness and ability to focus to explore things I always wanted to do but never thought I could do: be a more outdoors person and spend more time in nature (hiking, mountaineering), learn new skills (swimming), read more books (8 books and counting in 2021), teach at a university, but most importantly, find a purpose that puts everything in perspective and makes everything worthwhile. 

I was never good at sports growing up, however I vividly remember one winter afternoon while I was in college and playing cricket, I was hitting everything out of the park. I could not believe that it was me timing the ball to perfection, finding gaps in the field, and hitting fours and sixes. And then just like that, it was gone, I could never repeat what I did that one afternoon, until more than two decades later one morning in the gym when I was doing a circuit of heavy deadlifts. I experienced what people call the **flow** state, I was able to achieve a 1 rep max on the deadlift and then go on to do the same on the benchpress, my focus was razor sharp, I felt no fatigue, heart rate was normal and I felt I could go on and on (it is especially at such times you need a good trainer who takes an executive decision to say no, this is enough, we are stopping now). It was a magical feeling and unlike that one winter afternoon of decades past, I do not want it to be a moment in time. I want to keep experiencing it now, and decades from now.

## The road ahead, from the first edition

What do we do when our target is met? The proverbial last mile is extremely hard as we are finding out and you can see in the charts presented in the previous chapters. Going with the optimistic assumption that we will achieve our weight and strength targets in the next 2 to 3 months, what do we do next?

The natural answer is that sustaining ourselves at that level of health is going to be a challenge in itself and and a tough one at that. What I feel at this time is that this has been such a hard fought (not to mention expensive) battle that the very thought that this could all go away sends a chill down the spine. It is also true that I have been up this hill before and slided back.

What is going to help is to set a new target, that will change the problem from just being one of sustainence to one of achieving a new target. Setting your eyes on the next peak gives an aspirational goal to look up to. Aspirations make you wake up in the morning and get moving. When aspirations are standing on the legs of recent achievements, anything is possible.

My younger one wants to get into wrestling, he is 7 years old. Maybe he and I will train together some day! Let's see where this mid life "obsession" takes me.

## Parting thoughts...

What I realized over the past two years was that what most people perceive as a "boring & disciplined" life is actually a "healthy and therefore happy" life. I am a person of far less eloquence to say it better than this quote ([ascribed to Victor Frankel](https://www.univie.ac.at/logotherapy/quote_stimulus.html))

<i>Between stimulus and response lies a space. In that space lie our freedom and power to choose a response. In our response lies our growth and our happiness.</i>


